---
title: "Inventory Analysis Documentation"
author: "Anandhu Kannempilly Parambil"
date: "October 2026"
output: html_document
---

# AI Assistance Declaration

AI Assistance Declaration: I used ChatGPT (GPT-5.6 Sol, OpenAI, October 2026) for Markdown organization, R Markdown formatting suggestions, and writing assistance. Prompts used are documented below. I manually reviewed the documentation and compared the examples with my R script. All final calculations are done by myself. I am responsible for the accuracy and originality of this work.

# Purpose

The purpose of this R Markdown document is to explain the `inventory_analysis.R` script.

The script uses a synthetic inventory dataset to:

- Store product information.
- Calculate inventory value.
- Calculate total inventory value.
- Identify products below their reorder levels.
- Summarize inventory quantities by category.

No real or personal data is used.

# Inputs

The script uses a synthetic dataset containing:

- `Product_ID`
- `Product_Name`
- `Category`
- `Quantity`
- `Reorder_Level`
- `Unit_Price`

The inventory dataset is created using the following R code:

```{r}
inventory <- data.frame(
  Product_ID = c("P001", "P002", "P003", "P004", "P005",
                 "P006", "P007", "P008", "P009", "P010"),

  Product_Name = c(
    "Laptop", "Keyboard", "Office Chair", "Desk",
    "Printer Paper", "Pens", "Monitor",
    "Filing Cabinet", "Stapler", "Mouse"
  ),

  Category = c(
    "Electronics", "Electronics", "Furniture",
    "Furniture", "Office Supplies", "Office Supplies",
    "Electronics", "Furniture", "Office Supplies",
    "Electronics"
  ),

  Quantity = c(25, 8, 12, 4, 50, 18, 7, 9, 30, 14),

  Reorder_Level = c(10, 15, 5, 6, 20, 25, 10, 4, 10, 12),

  Unit_Price = c(850, 35, 180, 250, 8, 2, 220, 140, 12, 25)
)

inventory
```

# Processing

## Calculate Inventory Value

Inventory value is calculated by multiplying the quantity of each product by its unit price.

```{r}
inventory$Inventory_Value <-
  inventory$Quantity * inventory$Unit_Price

inventory
```

## Calculate Total Inventory Value

The `sum()` function calculates the total value of all inventory.

```{r}
total_inventory_value <- sum(inventory$Inventory_Value)

total_inventory_value
```

## Identify Low-Stock Products

A product is considered low stock when its current quantity is below its reorder level.

```{r}
low_stock <- inventory[
  inventory$Quantity < inventory$Reorder_Level,
]

low_stock
```

## Summarize Inventory by Category

The `aggregate()` function calculates total inventory quantity for each category.

```{r}
category_summary <- aggregate(
  Quantity ~ Category,
  data = inventory,
  FUN = sum
)

category_summary
```

# Outputs

The analysis produces:

1. The inventory dataset.
2. Inventory value for each product.
3. Total inventory value.
4. Products below their reorder levels.
5. Total quantity by product category.

# Expected Findings

The expected low-stock products are:

- Keyboard
- Desk
- Pens
- Monitor

The expected total inventory value is **$28,656**.

# Verification

I verified the project by comparing the documentation with the R script and manually checking selected calculations.

For example:

- Laptop inventory value: `25 × $850 = $21,250`
- Keyboard quantity is `8`, which is below its reorder level of `15`.

The R Markdown document should also be knitted or previewed in RStudio or Posit Cloud to verify that the code chunks and headings render correctly.

# Conclusion

This project demonstrates how R can be used for basic inventory analysis. The script calculates inventory values, identifies products that may need replenishment, and summarizes inventory by category.

R Markdown provides a clear way to combine explanations and R code in one document.

# AI Assistance

The following prompts were used with ChatGPT:

**Seed Prompt:**

> Here's my R script description. Suggest Markdown formatting and code block examples.

**Refinement Prompt:**

> Add syntax highlighting and improve section organization using Purpose, Inputs, and Outputs.

**Critique/Validation Prompt:**

> Is the Markdown consistent with R Markdown best practices?

I reviewed the AI-generated suggestions and compared them with the final R script before including them in the documentation.
